Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio

### 2017-05-14 eval.tex
space fixes

\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
	\end{minipage}
	\vspace{-6mm}
	\caption{\emph{Observed spectrum of television signal}. The left hand image shows rawspectrum of the TV signal with the centre frequency of \SI{626}{\mega\hertz}. The right hand image shows the smoothed spectrum, which is  normalized to maximum average. The average is  shown by the horizontal red line.}
	\label{fig:tv_record} 
		\vspace{-6mm}
\end{figure}

>>>>

\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
	\end{minipage}
	\vspace{-6mm}
	\caption{\emph{Observed spectrum of television signal}. The left hand image shows the raw spectrum of the TV signal, with centre frequency of \SI{626}{\mega\hertz}. The right hand image shows the smoothed spectrum normalized to maximum average. The average is  shown by the horizontal red line.}
	\label{fig:tv_record} 
		\vspace{-6mm}
\end{figure}

### 2017-05-14 introduction.tex
space fixes

Backscatter communication enables wireless transmissions
at energy consumption which is several orders of magnitude
lower than traditional radios. Backscatter achieves ultra-low
power wireless transmissions by reflecting or absorbing ambient
wireless signals, which as an operation consumes only 
\SI{}{\micro\watt}s of power consumption~\cite{liu_ambient_2013}. 
As a consequence backscatter
communication is emerging as the mechanism of choice to network
devices operating on harvested energy.  
Over the past few years there has been significant progress
to make backscatter a viable mechanism to network Internet of
Things~(IoT) devices. Traditional backscatter systems, like RFID readers, required a
dedicated device to generate the necessary carrier signal, which were later
reflected by the tags. On the other hand, state-of-the art systems 
does away the need for a dedicated device to generate carrier signals. 
Recent backscatter systems leverage already deployed infrastructure of
devices to generate carrier signal~\cite{varshney2016lorea,iyer2016inter}, or ambient WiFi~\cite{zhang2016hitchhike,kellogg2015wi} or TV
signals~\cite{liu_ambient_2013,parks_turbocharging_2014}.

>>>>

Backscatter communication enables wireless transmissions
at energy consumption which is several orders of magnitude
lower than traditional radios. Backscatter achieves ultra-low
power wireless transmissions by reflecting or absorbing ambient
wireless signals, which as an operation consumes only 
\SI{}{\micro\watt}s of power consumption~\cite{liu_ambient_2013}. 
As a consequence backscatter
communication is emerging as the mechanism of choice to network
devices operating on harvested energy.  
Over the past few years there has been significant progress
to make backscatter a viable mechanism to network Internet of
Things~(IoT) devices. Traditional backscatter systems, like RFID readers, required a
dedicated device to generate the necessary carrier signal
reflected by the tags. On the other hand, state-of-the art systems 
does away the need for a dedicated device to generate carrier signals. 
Recent backscatter systems leverage already deployed infrastructure of
devices to generate carrier signal~\cite{varshney2016lorea,iyer2016inter}, or ambient WiFi~\cite{zhang2016hitchhike,kellogg2015wi} or TV
signals~\cite{liu_ambient_2013,parks_turbocharging_2014}.

### 2017-05-14 paper.tex
space fixes

\section{Design}

In this section we describe the most important aspects of our design.

@@

Our backscatter transmitter is designed to be used with the television
signal, which has its center frequency at 626 MHz. To that end, we build
an
antenna using an unetched PCB plate as the ground plane and a wire that acts
as the vertical antenna element.
This antenna works well in the range from 626 to 245 MHz, which has been
measured with a vector network analyzer (VNA).

@@

\subsection{Receiver}
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{./fig/receiver_arch}
\caption{\emph{Design of the ambient backscatter receiver.} The flow of operation of the
receiver is from the left to right.}
\label{fig:receiver_arch}
\end{figure}
The receiver consists of three main modules interconnected by a bus. The
RTL-SDR block is in charge of collecting samples from the SDR-device. The
demodulator turns the sampled signal into a streams of ones and zeroes.
Finally, the Decoder block is in charge of detecting frames. We provide our C++
receiver code publicly available under~\cite{s3xm3x_backscatterBASKReceiver}.

@@

The resulting values are sent through the bus to the decoder. After the decoder
received the samples it decides when a frame is sent or when the channel is
idle. The decoder also includes a function to correlate the received bitstream
with an expected pattern. Hence it is able to determine the bit error ratio
(BER). Furthermore the code we have written so far subsumes different
simulators for e.g. playing back recorded data. Another handy utility, which
has been written in Octave, is a tool (an oscilloscope) to look at the data,
which is written into files by the demodulator. An example of a recorded data
stream with 25 kS/s can be seen in Figure \ref{fig:transmission}.  
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{./fig/transmission}
\caption{Start of a transmission of 101010.. with 1 kbit/s with idle state before. Data from the Octave oscilloscope. Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5. This is due to a low signal strength. Sampling frequency is 25 kHz.}
\label{fig:transmission}
\end{figure}

@@

\bibliographystyle{IEEEtran}
\bibliography{paper}

>>>>

\section{Design}
In this section, we describe the design 
of our system. We first describe the
design of the transmitter, next we describe 
the design of the receiver.

@@

We design our backscatter transmitter to be used with ambient
television signals.  We design the transmitter for ambient
television signal present in the band with  
center frequency of \SI{626}{\mega\hertz}. Crucial to the design
of the transmitter is the implementation of an antenna, as
the antennas are known to be  frequency selective. To this end, we design an
antenna on an unetched board made of FR4 substrate commonly
used to design PCBs. The board acts like the ground plane, at the center
of the board we have a wire acting as a monopole antenna element. 
We have verified using a vector network analyzer (VNA), that the antenna
can operate from frequency range most commonly
used for television broadcast, 
from \SI{245}{\mega\hertz} to \SI{626}{\mega\hertz}.

@@

\subsection{Receiver}
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{./fig/receiver_arch}
\vspace{-6mm}
\caption{\emph{Design of the ambient backscatter receiver.} The flow of operation of the
receiver is from the left to right.}
\label{fig:receiver_arch}
\vspace{-6mm}
\end{figure}
The receiver consists of three main modules interconnected by a bus. The
RTL-SDR block is in charge of collecting samples from the SDR-device. The
demodulator turns the sampled signal into a streams of ones and zeroes.
Finally, the Decoder block is in charge of detecting frames. We provide our C++
receiver code publicly available under~\cite{s3xm3x_backscatterBASKReceiver}.

@@

The resulting values are sent through the bus to the decoder. After the decoder
received the samples it decides when a frame is sent or when the channel is
idle. The decoder also includes a function to correlate the received bitstream
with an expected pattern. Hence it is able to determine the bit error ratio
(BER). Furthermore the code we have written so far subsumes different
simulators for e.g. playing back recorded data. Another handy utility, which
has been written in Octave, is a tool (an oscilloscope) to look at the data,
which is written into files by the demodulator. An example of a recorded data
stream with 25 kS/s can be seen in Figure \ref{fig:transmission}.

\begin{figure}[h]
\centering
\vspace{-30mm}

\includegraphics[width=\columnwidth]{./fig/transmission}
\vspace{-30mm}
\caption{Start of a transmission of 101010.. with 1 kbit/s with idle state before. Data from the Octave oscilloscope. Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5. This is due to a low signal strength. Sampling frequency is 25 kHz.}
\label{fig:transmission}
\vspace{-6mm}

\end{figure}

@@

\bibliographystyle{abbrv}
\bibliography{paper}
